Plots

### 2025-05-26 src/plots.Rmd
Add color and fix fig 1 brief methods v2

```{r, echo=FALSE, eval=FALSE}
#install.packages("openVA")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
#install.packages("glue")
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
library(glue)
```

@@

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = NULL,
        y_title = NULL,
        text_mid_size = 2 * text_scale,
        text_mid_vjust = 2.5,
        text_end_size = 3 * text_scale,
        text_hjust = 0.1,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 1.5,
        min_points_size = 2,
        max_points_size,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        ...
    ) {

@@

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_point( # mid points
            data = df_mids,
            size = mid_points_size
        ) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, " (", round(min_metric, 2), ")")
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(model_max, " (", round(max_metric, 2), ")")
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            panel.border = element_rect(
                color = "black",
                fill = NA,
                linewidth = 0.75
            ),
            legend.position = "none",
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Plot data
fig_pccc_csmf <- ggplot(
    df_plot,
    aes(
        x = metric_value,
        y = model_name,
        fill = metric
    )
) +
    geom_bar(
        stat = "identity",
        position = position_dodge(),
        color = "black",
        width = 0.7
    ) +
    geom_label(
        data = df_plot %>% filter(metric == "PCCC"),
        aes(
            label = metric_label,
            x = 0.085
        ),
        color = "black",
        fill = "white",
        size = 3,
        label.r = unit(0, "pt"),
        vjust = -0.5
    ) +
    geom_label(
        data = df_plot %>% filter(metric == "CSMF Accuracy"),
        aes(
            label = metric_label,
            x = 0.085
        ),
        color = "black",
        fill = "white",
        size = 3,
        label.r = unit(0, "pt"),
        vjust = 1.5
    ) +
    scale_fill_manual(
        values = c("black", "gray"),
        guide = guide_legend(reverse = TRUE)
    ) +
    scale_x_continuous(
        limits = c(0, 1),
        breaks = seq(0, 1, 0.1)
    ) +
    labs(
        title = NULL,
        x = NULL,
        y = NULL
    ) +
    theme_minimal() +
    theme(
        panel.border = element_rect(
            color = "black",
            fill = NA,
            linewidth = 0.75
        ),
        legend.position = c(.875, .5),
        legend.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 10),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8)
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ", ", round(perc_adult), "%)"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ", ", round(perc_child), "%)"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ", ", round(perc_neo), "%)"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_title = "PCCC",
    text_mid_vjust = 3,
    sort_group = F,
    x_limits = c(0.2, 0.8),
    x_expand = waiver()
)

>>>>

```{r, echo=FALSE, eval=FALSE}
#install.packages("openVA")
#install.packages("tidyverse")
#install.packages("dplyr")
#install.packages("ggrepel")
#install.packages("patchwork")
#install.packages("glue")
#install.packages("kableExtra)
```

@@

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
library(glue)
library(kableExtra)
```

@@

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = NULL,
        y_title = NULL,
        text_mid_size = 2 * text_scale,
        text_mid_vjust = 2.5,
        text_end_size = 3 * text_scale,
        text_hjust = 0.1,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 2.5,
        min_points_size = 2,
        max_points_size,
        x_limits = c(0, 1),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -2)),
        ...
    ) {

@@

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_point( # mid points
            aes(color = .data[[color]]),
            size = mid_points_size
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, " (", round(min_metric, 2), ")")
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(model_max, " (", round(max_metric, 2), ")")
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            panel.border = element_rect(
                color = "black",
                fill = NA,
                linewidth = 0.75
            ),
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Plot data
fig_pccc_csmf <- ggplot(
    df_plot,
    aes(
        x = metric_value,
        y = model_name,
        fill = metric
    )
) +
    geom_bar(
        stat = "identity",
        position = position_dodge(),
        color = "black",
        width = 0.7
    ) +
    geom_label(
        data = df_plot %>% filter(metric == "PCCC"),
        aes(
            label = metric_label,
            x = 0.085
        ),
        color = "black",
        fill = "white",
        size = 3,
        label.r = unit(0, "pt"),
        vjust = -0.5
    ) +
    geom_label(
        data = df_plot %>% filter(metric == "CSMF Accuracy"),
        aes(
            label = metric_label,
            x = 0.085
        ),
        color = "black",
        fill = "white",
        size = 3,
        label.r = unit(0, "pt"),
        vjust = 1.5
    ) +
    scale_x_continuous(
        limits = c(0, 1),
        breaks = seq(0, 1, 0.1)
    ) +
    labs(
        title = NULL,
        x = NULL,
        y = NULL
    ) +
    theme_minimal() +
    theme(
        panel.border = element_rect(
            color = "black",
            fill = NA,
            linewidth = 0.75
        ),
        legend.position = c(.875, .5),
        legend.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 10),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8)
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ", ", round(perc_adult), "%)"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ", ", round(perc_child), "%)"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ", ", round(perc_neo), "%)"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    x_title = "PCCC",
    text_mid_vjust = 3,
    sort_group = F,
    x_limits = c(0.2, 0.8),
    legend_margin = margin(-8, -2, -2, -2),
    x_expand = waiver()
)
